EvoSwarm: A Cloud-Native Architecture for the Reproducible Execution of Multi-population Optimization Algorithms

### 2020-05-16 draft.tex
Update draft.tex

% Mention cost here - JJ

@@

\subsubsection{Setup} % Again, this in the index with no context should say something. Please flesh it out.

To answer the first research question, we need to evaluate the effect of these parameters on the scalability of the system,   we
propose  an experiment in which we selected $f_4$ (Skew Rastrigin-Bueche separable) from the
BBOB testbed. This function has been used because it is computationally demanding,
and in higher dimensions has been difficult for PSO \cite{el2009black} and GAs
\cite{nicolau2009application} to solve with the required FEs. As we
are only comparing in terms of execution time, the results from a single function
can be better understood, since there are fewer factors involved.
Following the procedure described in the previous section, we executed 
two sets of experiments with five and ten initial populations,
repeating each experiment 30 times using  1, 2, 4, and 8 workers.  We kept the same 
maximum number of FEs, changing the relevant parameters for this. 
See Tables \ref{tab:params:10} and \ref{tab:params:5} 
for the complete list. 
%
\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{time_diff}
  \caption{Boxplot of the median time required for the execution of 30 
  instances of function $f_4$, in seconds, different number of workers and populations sizes.  }
  \label{fig:spworker:time}
\end{figure}
%
\begin{table}[]
  \caption{The median of the time required to finish an instance of the function $f_4$, 30 runs
  with different number of workers and populations} % time or rate? - JJ
  \label{tab:time}
  \vspace{0.25cm}
    \centering
    \begin{tabular}{ccrr}
    \hline
              &         & 5 Populations    & 10 Populations \\
     Dimension& Workers & Median of rate & Median of rate  \\
    \hline
          2   & 1w      & 2.96   & 3.15   \\
              & 2w      & 1.84   & 1.72   \\
              & 4w      & 1.22   & \textbf{0.96}   \\
              & 8w      & 1.05   & \textbf{0.59}   \\
    \hline
          3   & 1w      & 6.53   & \textbf{7.76}   \\
              & 2w      & 3.64   & 3.83   \\
              & 4w      & 2.33   & 2.26   \\
              & 8w      & 2.17   &  \textbf{1.36}   \\
    \hline
          5   & 1w      & \textbf{12.33}  & 14.19  \\
              & 2w      & 6.63   & \textbf{7.41}   \\
              & 4w      & 4.25   & 4.34   \\
              & 8w      & 3.67   &  \textbf{2.41}   \\
    \hline
          10  & 1w      & \textbf{31.50}  & 38.30  \\
              & 2w      & \textbf{16.07}  & 19.31  \\
              & 4w      &  9.99   & 10.35  \\
              & 8w      & 8.97   &  \textbf{5.92}  \\
    \hline
          20  & 1w      &  116.35 & 124.51 \\
              & 2w      & \textbf{51.18}  & 63.68  \\
              & 4w      & 32.26  & 33.07  \\
              & 8w      & 28.64  & \textbf{18.25}  \\
    \hline
          40  & 1w      & \textbf{369.89} & 448.89 \\
              & 2w      & \textbf{198.43} & 234.50 \\
              & 4w      & 122.39 &  120.62 \\
              & 8w      & 122.31 & \textbf{63.35} \\
    \hline
  \end{tabular}
\end{table}

@@

In this experiment, we wanted to test how the number of workers affected the
time needed to reach a solution and the speedup that each configuration can
produce. Adding more populations might add more overhead, so this might have a
noticeable effect. In principle, the main issue is how to find the right number
of workers for a particular problem, but in general, we wanted to check the
communication overhead when the size of the problem increases and how the
presented framework reacted to it.    % Added some explanation % here - JJ
We report the comparison regarding the execution time in Section~\ref{sec:exec-time}
and the summary of the speedup in section Section~ref{sec:speedup}.

\subsubsection{Execution Time}
\label{sec:exec-time}

@@

%
%
\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{evalspersecond}
  \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }
  \label{fig:spworker:evalspersecond}
\end{figure}
%
\subsubsection{Speedup}
\label{sec:speedup}

% Introduction saying why we are measuring this, how it's related to the objectives of the paper - JJ
To measure the speedup obtained by the addition of workers, we will use the
number of evaluations per second, instead of the time required to finish an
instance of a problem. The reason for this is that in some cases reaching the
desired target can require less work. We are going to measure the total increase
of work done in a second by adding more workers. The baseline for the comparison
will be the algorithm with a single worker, and the minimum amount of work available will be five
populations. This number is selected because the controller needs at least
three populations to migrate individuals between them, and we add two more to
keep even a single worker busy. Figure \ref{fig:spworker:evalspersecond} 
shows the number of function evaluation per second, achieved by a certain number of
workers, for every problem dimension, for five and ten dimensions of $f_4$.  % Why did you choose f4? - JJ

@@

Table~\ref{tab:speedup:test} shows the speedup obtained taking as the baseline
the previous configuration, for example, when increasing from four to eight
workers, and the p-rate for the left-tailed Wilcoxon test. There was an
increment in all cases, but again, for smaller dimensions, five populations with
fewer workers is a better option (see values underlined in the table) with the best speedup, achieved for 40 dimensions, shown in bold. The worst speed was going from four to eight workers 
at higher dimensions (underlined in the table). For ten populations, overall 
has the linear speedup % don't understand this - JJ
but the best is from four to eight workers at 40 dimensions
(shown in bold).

% No idea where this fits in. I think it's the second question in 7, 
% so it should probably be a subsection - JJ

@@

% Write an introduction - JJ

@@

\subsubsection{Setup}

@@

Therefore, in this experiment we compare a multi-population with  GA- and PSO- only populations,
versus an ensemble of GA and PSO algorithms. We have tested on the first five functions of the
BBOB testbed; we have used ten populations and eight workers for the experiment and the
same parameters as before.

>>>>

% Mention cost here - JJ
% I putted earlier literally at the end of the section :) - M

@@

From these parameters, we give more importance in this paper to the number of
populations, because they are the basic unit of work that can be executed in
parallel. Together with the size of each population and the number of
generations, this parameter determines the maximum number of function
evaluations for a single iteration. We do not have control over each algorithm's
complexity, but future analyses could focus on balancing populations according
to this.

In this experiment, we evaluate how the number of populations and the number of
workers in the system are related to the scalability of the system. In the next
section, we describe the experimental setup, and a discussion of the results
follows, with respect to execution time (Section~\ref{sec:exec-time}) 
and speedup Section~\ref{sec:speedup}.

\subsubsection{Setup} % Again, this in the index with no context should say something. Please flesh it out.
 % I will add more context in the parent section, and continue here

To answer the first research question, we need to evaluate the effect of 
the number of populations and the number of
workers on the scalability of the system, we 
propose  an experiment in which we selected $f_4$ (Skew Rastrigin-Bueche separable) from the
BBOB testbed. This function has been used because it is computationally demanding,
and in higher dimensions has been difficult for PSO \cite{el2009black} and GAs
\cite{nicolau2009application} to solve with the required FEs. As we
are only comparing in terms of scalability, the results from a single function
can be better understood, since there are fewer factors involved.
Following the procedure described in Section~\ref{setup}, we executed 
two sets of experiments with five and ten initial populations,
repeating each experiment 30 times using  1, 2, 4, and 8 workers.  We kept the same 
maximum number of FEs, changing the relevant parameters for this. 
See Tables \ref{tab:params:10} and \ref{tab:params:5} 
for the complete list. 
%
\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{time_diff}
  \caption{Boxplot of the median time required for the execution of 30 
  instances of function $f_4$, in seconds, different number of workers and populations sizes.  }
  \label{fig:spworker:time}
\end{figure}
%
\begin{table}[]
  \caption{The median of the time required to finish an instance of the function $f_4$, 30 runs
  with different number of workers and populations} % time or rate? - JJ
  \label{tab:time}
  \vspace{0.25cm}
    \centering
    \begin{tabular}{ccrr}
    \hline
              &         & 5 Populations    & 10 Populations \\
     Dimension& Workers & Median of rate & Median of rate  \\
    \hline
          2   & 1w      & 2.96   & 3.15   \\
              & 2w      & 1.84   & 1.72   \\
              & 4w      & 1.22   & \textbf{0.96}   \\
              & 8w      & 1.05   & \textbf{0.59}   \\
    \hline
          3   & 1w      & 6.53   & \textbf{7.76}   \\
              & 2w      & 3.64   & 3.83   \\
              & 4w      & 2.33   & 2.26   \\
              & 8w      & 2.17   &  \textbf{1.36}   \\
    \hline
          5   & 1w      & \textbf{12.33}  & 14.19  \\
              & 2w      & 6.63   & \textbf{7.41}   \\
              & 4w      & 4.25   & 4.34   \\
              & 8w      & 3.67   &  \textbf{2.41}   \\
    \hline
          10  & 1w      & \textbf{31.50}  & 38.30  \\
              & 2w      & \textbf{16.07}  & 19.31  \\
              & 4w      &  9.99   & 10.35  \\
              & 8w      & 8.97   &  \textbf{5.92}  \\
    \hline
          20  & 1w      &  116.35 & 124.51 \\
              & 2w      & \textbf{51.18}  & 63.68  \\
              & 4w      & 32.26  & 33.07  \\
              & 8w      & 28.64  & \textbf{18.25}  \\
    \hline
          40  & 1w      & \textbf{369.89} & 448.89 \\
              & 2w      & \textbf{198.43} & 234.50 \\
              & 4w      & 122.39 &  120.62 \\
              & 8w      & 122.31 & \textbf{63.35} \\
    \hline
  \end{tabular}
\end{table}

@@

% Added some explanation % here - JJ

%We report the comparison regarding the execution time in Section~\ref{sec:exec-time}
%and the summary of the speedup in section Section~ref{sec:speedup}.
% Move this to parent - M

\subsubsection{Results Regarding Execution Time}
\label{sec:exec-time}

% I think this is now a bit redundant or we need to move it to the parent section - M
% But is a good intro to the results.

In this experiment, we wanted to test how the number of workers affected the
time needed to reach a solution and the speedup that each configuration can
produce. Adding more populations might add more overhead, so this might have a
noticeable effect.

%In principle, the main issue is how to find the right number
%of workers for a particular problem, but in general, we wanted to check the
%communication overhead when the size of the problem increases and how the
%presented framework reacted to it.
% Commenting because we did not checked the communication over head - M

@@

%
%
\begin{figure}[h!tbp]
  \centering
  \includegraphics[width=\textwidth]{evalspersecond}
  \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }
  \label{fig:spworker:evalspersecond}
\end{figure}
%
\subsubsection{Results Regarding Speedup}
\label{sec:speedup}

% Introduction saying why we are measuring this, how it's related to the objectives of the paper - JJ
% This is done on the parent section, I renamed the sections to clarify this - M
To measure the speedup obtained by the addition of workers, we will use the
number of evaluations per second, instead of the time required to finish an
instance of a problem. The reason for this is that in some cases reaching the
desired target can require less work. We are going to measure the total increase
of work done in a second by adding more workers. The baseline for the comparison
will be the algorithm with a single worker, and the minimum amount of work available will be five
populations. This number is selected because the controller needs at least
three populations to migrate individuals between them, and we add two more to
keep even a single worker busy. Figure \ref{fig:spworker:evalspersecond} 
shows the number of function evaluation per second, achieved by a certain number of
workers, for every problem dimension, for five and ten dimensions of $f_4$.  % Why did you choose f4? - JJ This is mentioned on the setup - M

@@

Table~\ref{tab:speedup:test} shows the speedup obtained taking as the baseline
the previous configuration, for example, when increasing from four to eight
workers, and the p-rate for the left-tailed Wilcoxon test. There was an
increment in all cases, but again, for smaller dimensions, five populations with
fewer workers is a better option (see values underlined in the table) with the best speedup, achieved for 40 dimensions, shown in bold. The worst speed was going from four to eight workers 
at higher dimensions (underlined in the table). 
For ten populations, adding more workers resulted in a linear speedup, % don't understand this - JJ % Me either - M, rewritten  
but the best speedup is obtained when increasing from four to eight workers at 40 dimensions
(shown in bold).

% No idea where this fits in. I think it's the second question in 7, 
% so it should probably be a subsection - JJ
% done - M

@@

% Write an introduction - JJ
% I removed the Setup section, because we use the same configuration as before
% I will mention something about it - M
% mm but I don't like to have a single subsection, as I remember is not right.

@@

%\subsubsection{Setup}

@@

Therefore, in this experiment we compare a multi-population with GA and PSO only populations,
versus an ensemble of GA and PSO algorithms. We have tested on the first five functions of the
BBOB testbed; we have used ten populations and eight workers for the experiment and the
same parameters as before.

In order to run this experiment, we just needed to edit a configuration file and run 
the experiments. The plots presented in the results are generated by the BBOB framework
from the JOSN log file of each run. There was no additional setup needed.
